---
title: "Sanity-check NestedVar function"
output: 
    html_document:
        toc: true
author: Joyce Hsiao
date: "2015-10-29"
---


## Objective

Raunak Mundada working with Monnie McGee ran into an error message while simulating cell populations....


## Raunak's code

Copied from our correspondence.

```{r, eval = FALSE}
> rm(list=ls(all=T))

> source('F:/SMU_Courses/ResearchWork/flowMap_1.6.0/flowCytometry/simulateFCM_helpers.R')

> source('F:/SMU_Courses/ResearchWork/flowMap_1.6.0/flowCytometry/simulateFCM.R')

> library(parallel)

> library(doParallel)

> library(flowMap)

> ref_id <- 7 # id of the reference sample

> sim_id <- 8 # id of the cell population that is to be simulated

> sam1 <- realData #read from sample7.txt

> temp <- make.propcase(id = sim_id,simulatedata = realData,sampleMethod = "proportional",samToMatch = 3)
```



## Trial run

Load packages.

```{r}
require(flowMap)
require(sn)
require(MBESS)
```

Import helper functions.

```{r}
dir <- "~/Dropbox/GitHub/simulateFlow/"
source(file.path(dir, "simulateFCM_helpers.R"))
```

Import the real FCM data used in the paper for simulation.

```{r}
## Import a real flow cytometry data set
realData <- read.table(file.path(dir, "sample7.txt"), sep="\t", header=TRUE)
```

Include 4 marker channels: CD14, CD23, CD3, and CD19, as well as the cell population identifying label

```{r}
realData <- realData[,c(1:4,6)]
colnames(realData)[5] <- "id"
```

Recode cell population ID numbers to a sequence of  1 to 9

```{r}
iiChange <- realData$id > 5
realData$id[iiChange] <- realData$id[iiChange] - 1
```

Parameter settings (see Supplemental information)

```{r}
## Specificy mean and covariance parameters of each 
## cell populations. These parameters are also described
## in the paper. 

# CP1 covariance
covarr_pop1 <- .8*cov( realData[realData$id==1,1:4] )

# CP2 covariance
covarr_pop2 <- matrix(c(3500,0,0,0,0,18000,0,0,0,0,
                        6100,0,0,0,0,1200), 
                      nrow = 4, ncol = 4, byrow = FALSE)

# CP3 covariance
rho12 = 0.9
rho13 = 0.9
rho14 = 0.2
rho23 = 0.7
rho24 = 0.2
rho34 = 0.2
corrmat <- matrix(c(1,rho12,rho13,rho14,rho12,1,rho23,
	rho24,rho13,rho23,1,rho34,rho14,rho24,rho34,1), 4, 4)
covarr_pop3 <- cor2cov(corrmat, sd=.4*c(sqrt(3862),
		sqrt(2627),sqrt(4672),sqrt(9948)))

# CP4 covariance
covarr_pop4 <- cov( realData[realData$id==4,1:4] )

# CP5 covariance
rho12 = -0.1
rho13 = -.2
rho14 = .1
rho23 = 0.2
rho24 = -0.2
rho34 = -0.2
corrmat <- matrix(c(1,rho12,rho13,rho14,rho12,1,rho23,
                    rho24,rho13,rho23,1,rho34,rho14,rho24,rho34,1),4,4)
covarr_pop5 <- cor2cov(corrmat, sd=.4*c(sqrt(2300),sqrt(2700),
                                        sqrt(2300),sqrt(2600)))
# CP6 covariance
covarr_pop6 <- cov( realData[realData$id==6,1:4] )

# CP7 covariance
covarr_pop7 <- cov( realData[realData$id==7,1:4] )

# CP8 covariance
rho12 = .6
rho13 = .6
rho14 = .3
rho23 = .6
rho24 = .3
rho34 = .3
corrmat <- matrix(c(1,rho12,rho13,rho14,rho12,1,rho23,
	rho24,rho13,rho23,1,rho34,rho14,rho24,rho34,1),4,4)
covarr_pop8 <- cor2cov(corrmat, sd=.4*c(sqrt(2440),sqrt(2080),
	sqrt(2552),sqrt(1606)))

# CP9 covariance
rho12 = .6
rho13 = .5
rho14 = .5
rho23 = .7
rho24 = .4
rho34 = .4
corrmat <- matrix(c(1,rho12,rho13,rho14,rho12,1,rho23,
                    rho24,rho13,rho23,1,rho34,rho14,rho24,rho34,1),4,4)
covarr_pop9 <- cor2cov(corrmat, sd=.4*c(sqrt(9444),
                                        sqrt(12856),sqrt(9483),sqrt(5423)))


## Coerce the cell population paramters of mean, covariance, and skewness 
## into a list object 

CP1 <- list(xi = c(113,129,190,300),
		Omega = covarr_pop1,
		alpha = c(0,0,0,-4),df = 1 )
CP2 <- list(xi = c(68,585,95,497), 
		Omega = covarr_pop2, 
		alpha = c(0,-2.5,0,0), df = 3 )
CP3 <- list(xi = c(392,464,416,563), 
            Omega = covarr_pop3, 
            alpha = c(0,0,0,0), df = 3 )
CP4 <- list(xi = c(214,277,270,240), 
		Omega = .7*covarr_pop4, 
		alpha = c(0,0,-7,0), df = 3 )
CP5 <- list(xi = c(186,248,252,276), 
            Omega = covarr_pop5, 
            alpha = c(0,0,0,3), df = 2.5 )
CP6 <- list(xi = c(670,191,218,262), 
		Omega = .8*covarr_pop6, 
		alpha = c(-6,0,0,1.5), df = 3 )
CP7 <- list(xi = c(98,157,650,260), 
		Omega = .5*covarr_pop7, 
		alpha = c(0,-1,-8,-2), df = 3 )
CP8 <- list(xi = c(238,303,290,293), 
		Omega = covarr_pop8, 
		alpha = c(3,3,2,2), df = 3 )
CP9 <- list(xi = c(530,592,496,412), 
            Omega = 2*covarr_pop9, 
            alpha = c(0,0,0,0), df = 4 )

paramlist <- list(CP1,CP2,CP3,CP4,CP5,CP6,CP7,CP8,CP9)
```

*Beginning of Raunak's test.

Select the populations of interest.

```{r}
ref_id <- 7 # id of the reference sample
sim_id <- 8 # id of the cell population that is to be simulated
sam1 <- realData #read from sample7.txt
```


```{r}
temp <- make.propcase(id = sim_id,
                      simulatedata = realData,
                      sampleMethod = "proportional",
                      samToMatch = 3)
```

So the debugging begins...

```{r}
id = sim_id
simulatedata = realData
sampleMethod = "proportional"
samToMatch = 3

# Specificy proportions of interest
prop_list <- c(1, 10, seq(25, 150, 25) )/ 100

# Convert proportions to counts
simprops <- round( 100*table(simulatedata$id)[id]*prop_list/
                     length(simulatedata$id[id]), 2)

propcase <- simSam2(pop = id, param = simprops/100, sam1=simulatedata,
                    xi = paramlist[[id]]$xi, Omega = paramlist[[id]]$Omega, 
                    alpha = paramlist[[id]]$alpha, df = paramlist[[id]]$df) 
```

Debug simSam2().

```{r}
pop = id
param = simprops/100
sam1 = simulatedata
xi = paramlist[[id]]$xi
Omega = paramlist[[id]]$Omega
alpha = paramlist[[id]]$alpha
df = paramlist[[id]]$df

#simSam2 <- function(pop,param,sam1,xi,Omega,alpha,df) {
  
  tempmat <- sam1[sam1$id!=pop,]
  nn <- nrow(sam1)
  
  sam2 <- lapply(1:length(param),function(i) {

    nsam <- round(param[i])	
    if (nsam==0) {
      tempmat
    } else if (nsam > 0) {
      mat <- simOnePop(n=nsam, xi=xi, Omega=Omega, 
                       alpha=alpha, df=df)
#       mat$id <- rep(pop, nsam)
#       colnames(mat)[1:4] <- colnames(tempmat)[1:4]
#       rbind(mat, tempmat)
    }
  })
  # return(sam2)
# }
```

Debug simOnePop()

```{r}
n = nsam
nn = n
#simOnePop <- function(nn,xi,Omega,alpha,df) {
  
rsam <- rmst_ed(n = nn,xi = xi,Omega=Omega,alpha=alpha, df=df)
# IIout <- rowSums(rsam<0 | rsam>1000) >0
# rsam <- rsam[!IIout,]
# nn <- sum(IIout)
#   while(nn > 0) {
#     temp <- rmst_ed(n=nn,xi = xi,Omega=Omega,alpha=alpha,df=df)
#     IIout <- rowSums(temp<0 | temp>1000) >0
#     rsam2 <- temp[!IIout,]
#     nn <- sum(IIout)
#     rsam <- rbind(rsam,rsam2)
#   }
#   rsam <- data.frame(rsam,check.names=F,row.names=NULL)
#   return(rsam)	
# }

```


Debug rmst_ed().

```{r}
n = nn
xi = xi
Omega = Omega
alpha = alpha
df=df
dp = NULL
# rmst_ed <- function(n = 1, 
#                     xi=rep(0,length(alpha)), 
#                     Omega, alpha, df=Inf, dp=NULL)
# { 
#   if(!(missing(alpha) & missing(Omega)) && !is.null(dp)) 
#     stop("You cannot set both component parameters and dp")
#   if(!is.null(dp)){
#     if(!is.null(dp$xi)) xi <- dp$xi
#     else
#       if(!is.null(dp$beta)) xi <- as.vector(dp$beta)
#       Omega <- dp$Omega
#       alpha <- dp$alpha
#       df <- dp$df
#   }  
  d <- length(alpha)
  x <- if(df==Inf) 1 else rchisq(n,df)/df
  z <- rmsn(n, rep(0,d), Omega, alpha)
  # y <- sweep(z/sqrt(x),2, STAT=xi, FUN="+")
  # y <- t(xi+ t(z/sqrt(x)))
  # attr(y,"parameters") <- list(xi=xi, Omega=Omega, alpha=alpha, df=df)
  # return(y)
# }
```


Debug rmsn()

```{r}
n = n
x = rep(0,d)
tau = 0
dp = NULL
# function (n = 1, xi = rep(0, length(alpha)), Omega, alpha, tau = 0, 
#     dp = NULL) 
# {
#     if (!is.null(dp)) {
#         dp0 <- dp
#         dp0$nu <- NULL
#         if (is.null(dp0$tau)) 
#             dp0$tau <- 0
#     }
#    else 
      dp0 <- list(xi = xi, Omega = Omega, alpha = alpha, tau = tau)

    if (any(abs(dp0$alpha) == Inf)) 
#         stop("Inf's in alpha are not allowed")
    lot <- dp2cpMv(dp = dp0, family = "SN", aux = TRUE)
    d <- length(dp0$alpha)
    y <- matrix(rnorm(n * d), n, d) %*% chol(lot$aux$Psi)
    if (dp0$tau == 0) 
        truncN <- abs(rnorm(n))
#    else truncN <- qnorm(runif(n, min = pnorm(-dp0$tau), max = 1))
    truncN <- matrix(rep(truncN, d), ncol = d)
    delta <- lot$aux$delta
    z <- delta * t(truncN) + sqrt(1 - delta^2) * t(y)
    y <- t(dp0$xi + lot$aux$omega * z)
    attr(y, "family") <- "SN"
    attr(y, "parameters") <- dp0
    return(y)
}
```


## Session information

```{r}
sessionInfo()
```

